---
title: "Assignment 2"
author: "Niels Krarup"
date: '2022-05-07'
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, echo=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)

```

## Exercise 4.b - Finding minimum by numeric optimization

To find the minimum we use the stats::optim function from default R.

```{r, fig.align='center', fig.width=12}
fooX <- function(x) ((x-1)^2 + (x-3)^4 + 1)
fooY <- function(y) (2 + y^2)
#num
optX <- optim(2, fooX, method = "Brent", lower = 0, upper = 10)
optY <- optim(2, fooY, method = "Brent", lower = 0, upper = 10)

plot(fooX, xlim = c(-2,4), ylim = c(1,6), ylab = "EZ;do(.)")
points(optX$par, optX$value)
abline(v = optX$par, lty = 2)


curve(fooY, col = 2, add = T)
points(optY$par, optY$value, col = 2)
abline(v = optY$par, lty = 2, col = 2)

legend("topright", c("Do(X:=x)", "Do(Y := y)"), col = 1:2, lty = 1)
```


## Looking at solutions.

expected value of Z given X intervention
```{r}
optX[c("par", "value")]
```
expected value of Z given Y intervention
```{r}
optY[c("par", "value")]
```
it is seen that the minimum value is attained by intervening on $Y$ with at $do(Y := 0)$ with a corresponding expected value of $Z$ of $2$.

# b)
```{r, fig.align='center', fig.width=12}
foo_cf_X <- function(x) (x-3)^4 + (x-1)^2 -8*(x-3)^2 + 16
foo_cf_Y <- function(y) y^2 + 4
#num

plot(foo_cf_X, xlim = c(-2, 6), ylim = c(0,30), ylab = "EZ|(X,Y,Z) = (3,-4,20);do(.)")
curve(foo_cf_Y, col = 2, add = T)
abline(h = min(optX$value, optY$value), lty = 2)
legend("topleft", c("Z|(X,Y,Z) = (3,-4,20);do(X=x)", "Z|(X,Y,Z) = (3,-4,20);do(Y=Y)"), col = 1:2, lty = 1)

```

It can be seen from above that the lowest expected value of Z i.e. 2, which was from the intervention $do(Y:=0)$ is only possible to attai by interveening on $X$ given the observations.

We solve to finde these values

```{r}
r1 <- uniroot(f = function(x)foo_cf_X(x = x) - 2, lower = 0, upper = 1)
r2 <- uniroot(f = function(x)foo_cf_X(x = x) - 2, lower = 1, upper = 4)
plot(foo_cf_X, xlim = c(-2, 6), ylim = c(0,30), ylab = "EZ|(X,Y,Z) = (3,-4,20);do(.)")
abline(h = 2, lty = 2)
abline(v = c(r1$root, r2$root), col = 3, lty = 3)
```

the interventions on $X$ with counterfactuals equal to smallest expected value of Z are: 
`r round(c(r1$root, r2$root),4) `

# end of document